---
title: "Dirty Data Project - Task Three - Mahri"
output:
  html_document:
    df_print: paged
---

Seabird sightings from ships were recorded over many years at different latitudes.
Whether spotted as a solitary bird or as a large group, their numbers were 
recorded on each spotting along with their commonly known name, species specific
name and a name abbreviation. 
If details of a birds appearance (i.e. plumage phase), sex, or age were obvious 
to the spotter, this was noted as well. 

Having completed the questions, I now note that I should have separated details
such as age/plumage from each name as these birds are still the same species. 
This is something to come back to. 

```{r}
library(tidyverse)
library(readxl)
library(here)
library(stringr)
```


Though there were four excel sheets:
ship_data_by_record
bird_data_by_record
ship_data_codes
bird_data_codes 
The 3rd and 4th were only used to identify which columns may be of interest in 
the first two sheets, which were used for cleaning and analysis. 

There was no meta data to be removed.

The sheets were cleaned in an R Script and written to a csv file which has been 
read in here, below. 

```{r}
birds_and_ships_clean <- read_csv("../clean_data/birds_and_ships_clean.csv")
```

To get a quick idea of the variables and their types, glimps and names were used.
* Note: both have RECORD and RECORD ID so plan to match on Record ID - lots in 
  the birds birds data match with 1 in ship.
  
```{r}
glimpse(ship_data_by_record)
# 27 cols, 12,310 rows 

glimpse(bird_data_by_record)
# 26 cols, 49,019 rows

names(ship_data_by_record)
names(bird_data_by_record)
```
#### Cleaning

Janitor was used for an initial column clean. Columns were then renamed with the 
purpose of making them easier to read and those that were relevant to the 
questions were selected:

```{r}
library(janitor)

ship_data_records <- janitor::clean_names(ship_data_by_record)
bird_data_records <- janitor::clean_names(bird_data_by_record)

bird_data_renamed <- bird_data_records %>%
  rename("species_common_name" = "species_common_name_taxon_age_sex_plumage_phase",
         "species_scientific_name" = "species_scientific_name_taxon_age_sex_plumage_phase",
         "birds_counted" = "count") %>% 
  select(record, 
        record_id, 
        species_common_name, 
        species_scientific_name, 
        species_abbreviation, 
        birds_counted)

bird_data_renamed

ship_data_renamed <- ship_data_records %>% 
  rename("latitude" = "lat", 
         "longitude" = "long") %>% 
  select(record, 
         record_id, 
         date,
         latitude, 
         longitude)

ship_data_renamed
```


## Left Join of the two sheets. 

Joining on record_id, column order was rearranged, and I kept record numbers for 
both the ship and the birds but named both more appropriately for understanding.

```{r}
birds_and_ships_joined <- left_join(bird_data_renamed, 
                                    ship_data_renamed, by = "record_id") %>% 
  rename("record_num_birds" = record.x, 
         "record_num_ship" = record.y)  %>% 
  relocate(birds_counted, .before = 1) %>% 
  relocate(record_num_birds, .before = 7)

birds_and_ships_joined
```

## Further Cleaning

* Dropping appropriate NAs with regards to the questions (i.e. where bird count
  is equal to NA (there were 2699 of them)).
* Note that there are some bird names (of all 3 types) as "NA" and some latitudes 
  (28) as "NA", but I'll wait and see if that's an issue later?.

```{r}

birds_and_ships_joined %>% 
  arrange(!is.na(birds_counted)) %>% 
  summarise(count = sum(is.na(birds_counted)))
# there are 2699 records where bird count = NA 

birds_and_ships <- birds_and_ships_joined %>% 
  drop_na(birds_counted)

birds_and_ships %>% 
  summarise(count = sum(is.na(birds_counted)))
# No NAs in bird count 

birds_and_ships

```

Checks ran to make sure that all the columns are the correct type:
* Note: there are 28 latitude failures and 29 longitude failures - these are 
  all NA values. 
* The latitude NA values all had quite a small bird count so I don't think it is 
  an issue for the latitude question.
* These small numbers are probably needed for the solo sightings question though 
  so dropping them was not considered. 
  
```{r}
glimpse(birds_and_ships)

library(assertr)
birds_and_ships %>% 
  verify(latitude >= -90 & latitude <= 90)
birds_and_ships %>% 
  verify(longitude >= -180 & longitude <= 180)
# there are 28 latitude failures and 29 longitude failures... 

birds_and_ships %>% 
  summarise(count = sum(is.na(latitude)))
# there are 28 NA latitude values 
birds_and_ships %>% 
  summarise(count = sum(is.na(longitude)))
# there are 29 NA longitude values

birds_and_ships %>% 
  arrange(!is.na(latitude))
# small numbers as note above
```


#### **Analysis questions**

For each question where we ask for bird names, give the bird’s common name, 
scientific name and species abbreviation.


**QUESTION ONE**
**Which bird had the most individual sightings?**

I'm not exactly sure what the question is asking for but here are two options: 

1. The species of bird that was spotted on the most occasions (1 or 
  more of them spotted at once) was:
  
  Common Name: Wandering albatross sensu lato PL2
  Scientific Name: Diomedea antipodensis / exulans PL2
  Species Abbreviation: DIOANTEXU PL2
  This species of bird was spotted on 3099 separate occasions
  
2. For sightings of lone birds, which was observed the most:  

 Common Name: Wandering albatross sensu lato PL3
 Scientific Name: Diomedea antipodensis / exulans PL3
 Species Abbreviation: DIOANTEXU PL3
 This species of bird was spotted on it's own on 1707 separate occasions
 
```{r}

# 1
birds_and_ships %>% 
  select(species_abbreviation, 
         species_common_name,
         species_scientific_name,
         birds_counted, 
         record_id) %>% 
  group_by(species_common_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
  
# 2
birds_and_ships %>% 
  select(species_abbreviation, 
         species_common_name,
         species_scientific_name,
         birds_counted, 
         record_id) %>% 
  filter(birds_counted == 1) %>% 
  group_by(species_common_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

```

**QUESTION TWO**
**Which bird had the highest total count?**

The bird with the highest total count (982,553) was:

* Common Name: Short-tailed shearwater
* Scientific Name: Puffinus tenuirostris
* Species Abbreviation: PUFTEN
* This bird had a total count of 982,553

```{r}
birds_and_ships %>% 
  group_by(species_common_name) %>% 
  summarise(n_birds = sum(birds_counted)) %>% 
  arrange(desc(n_birds))
```

**QUESTION THREE**
**Which bird had the highest total count above a latitude of -30?**

This bird was seen alone or in groups above a latitude of -30 on 29 occasions. 
An individual count of these birds (either alone or in groups) shows them to 
have the highest total count of 855 birds:

  * Common Name: Wedge-tailed shearwater.
  * Scientific Name: Puffinus pacificus.
  * Species Abbreviation: PUFPAC.

```{r}
birds_and_ships %>% 
  select(species_abbreviation, 
         species_common_name,
         species_scientific_name,
         birds_counted, 
         latitude,
         record_id) %>% 
  filter(latitude > -30) %>% 
  group_by(species_common_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))


birds_and_ships %>% 
  select(species_abbreviation, 
         species_common_name,
         species_scientific_name,
         birds_counted, 
         latitude,
         record_id) %>% 
  filter(latitude > -30) %>% 
  group_by(species_common_name) %>% 
  summarise(n_birds = sum(birds_counted)) %>% 
  arrange(desc(n_birds))

```

**QUESTION FOUR**
**How many different types of birds were only ever seen in groups of 1?**

* Birds were firstly grouped by their common name and a filter was added to 
  select only those who had a maximum of "1" in all of their "bird count" rows.
* Each species was collected together using distinct() and the length of this 
  response was noted 
  
**Answer**
86 species of bird were only ever seen in groups of 1.
However - I believe this number should be less as I have neglected to distinguish 
between ages/plumage stage of the birds (these are still the same species). With 
more time, I shall return to this. 

```{r}
single_sightings <- birds_and_ships %>% 
  group_by(species_common_name) %>% 
  filter((max(birds_counted)) == 1) %>% 
  distinct(species_common_name) %>% 
  pull() %>% 
  length()
single_sightings

```


**QUESTION FIVE** 
**How many penguins were seen? (Hint: there are many types of penguin)**

* Still working on this one, currently have the answer 10

```{r}


penguins <- birds_and_ships %>%
  filter(str_detect(species_common_name, "[pP]enguin")) %>%
  #summarise(penguin_count, n())
  # pull() %>% 
  length()
  
# penguin_pattern <- "[*][pP]enguin"

# penguins <- birds_and_ships %>% 
#  str_detect(penguin_pattern) %>% 
#  length()
#penguins
# the above gives 9 false and 1 true. and then a length of 10... there are 9

#penguins_seen <- birds_and_ships %>% 
#  select(birds_counted, species_common_name) %>% 
#  filter(species_common_name %in% c("Penguin", "penguin")) 


```

