---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
```

```{r}
candy_2015 <- read_excel("../raw_data/boing-boing-candy-2015.xlsx")
candy_2016 <- read_excel("../raw_data/boing-boing-candy-2016.xlsx")
candy_2017 <- read_excel("../raw_data/boing-boing-candy-2017.xlsx")

here::here()
```

From glimpse 
```{r}
head(candy_2015)
glimpse(candy_2016)
glimpse(candy_2017)
```



```{r}
library(janitor)

```

```{r}

janitor_candy_2015 <- janitor::clean_names(candy_2015)
janitor_candy_2015

janitor_candy_2016 <- janitor::clean_names(candy_2016)
janitor_candy_2016

janitor_candy_2017 <- janitor::clean_names(candy_2017)
janitor_candy_2017
```


just looking at who is reporting back about these ones... 
```{r}

janitor_candy_2017 %>% 
  select(q3_age, q2_gender, q6_independent_m_ms, q6_green_party_m_ms)
# all types of people
janitor_candy_2016 %>% 
  select(york_peppermint_patties_ignore)
#nobody 
```

**REMOVE AND RENAME** 

Step 1 - Remove and add for each year

- For each of the three years, I removed columns that weren't candy having 
 searched the internet to check on those I was not sure of. I also looked into
 responses for a few of the columns (e.g. different m&ms colors) to see if there
 were mostly joke responses/ any reason for me to join them together (I kept
 them as is excpet for changing some names to make it clearer to me what they 
 were - e.g. party bag of m&ms)
- I removed columns by index from last position to first so that I could check 
 that I didn't remove the wrong columns/ affect the order as I went.
- Also adding a "year" column and removing "timestamp" from 2015 and 2016 
 for referencing the correct years after binding the rows of all three together
 (see if can work out extracting year from timestamp and moving over later). 
 Adding this after so that column index isn't ruined.
- Added a personal ID number just to try different ways of adding/ mutating
 columns. (note - went to base R to do this for 2016 & 17)
- I probably didn't have to remove all of this (depending on question and outcome) 
but I like that it gives me less to work with.

Step 2 - RENAME FOR EACH YEAR

- Renaming columns so they match other years when binding rows. 
- Considered merging the 2015 and 2016 "anonymous brown globs etc." columns with 
 their "Mary Janes" columns because in 2017 it is recorded as "anon brown... aka 
 Mary Janes"). But from looking at individuals responses, they don't match up so 
 I'll leave as "anonymous_black_and_orange_wrapper" and "mary_janes".


2015 REMOVE AND ADD 

```{r}
# names(janitor_candy_2015)

col_removed_candy_2015 <- janitor_candy_2015 %>% 
  select(-c(116:124), -c(97:113), -c(93:95), -c(90, 91), 
         -c(peterson_brand_sidewalk_chalk, spotted_dick, mint_leaves, 
            joy_joy_mit_iodine, minibags_of_chips, lapel_pins, kale_smoothie, 
            hugs_actual_physical_hugs, heath_bar, healthy_fruit, 
            creepy_religious_comics_chick_tracts, broken_glow_stick, 
            glow_sticks, generic_brand_acetaminophen, dental_paraphenalia, 
            cash_or_other_forms_of_legal_tender,
            vials_of_pure_high_fructose_corn_syrup_for_main_lining_into_your_vein, 
            box_o_raisins, timestamp)) %>% 
  add_column(year = "2015", .before = 1) %>% 
  mutate(id_number = row_number(), .before = 2)


col_removed_candy_2015
#view(col_removed_candy_2015)


```



2015 RENAME
```{r}
# Looking at differences between responses for Mary Janes thoughts

# col_removed_candy_2015 %>% 
#  select(anonymous_brown_globs_that_come_in_black_and_orange_wrappers, mary_janes)

candy_2015_renamed <- col_removed_candy_2015 %>% 
  rename(age = how_old_are_you, 
         trick_or_treating = are_you_going_actually_going_trick_or_treating_yourself,
         anonymous_black_and_orange_wrapper = 
           anonymous_brown_globs_that_come_in_black_and_orange_wrappers, 
         brach_not_including_candy_corn = brach_products_not_including_candy_corn, 
         restaurant_candy = 
           candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants, 
         hersheys_dark_chocolate = dark_chocolate_hershey, 
         gummy_bears = gummy_bears_straight_up, 
         hersheys_kissables = hershey_s_kissables, 
         hersheys_milk_chocolate = hershey_s_milk_chocolate, 
         licorice_black = licorice, 
         reeses_peanut_butter_cups = reese_s_peanut_butter_cups, 
         toblerone = tolberone_something_or_other, 
         peanut_m_ms = peanut_m_m_s, 
         chick_o_stick = chick_o_sticks_we_don_t_know_what_that_is, 
         circus_peanuts = those_odd_marshmallow_circus_peanut_things, 
         sea_salt_chocolate = 
           sea_salt_flavored_stuff_probably_chocolate_since_this_is_the_it_flavor_of_the_year)

```


2016 - REMOVE AND ADD 

```{r}
col_removed_candy_2016 <- janitor_candy_2016 %>% 
  select(-c(104, 105, 107:123), 
         -c(vicodin, vials_of_pure_high_fructose_corn_syrup_for_main_lining_into_your_vein, 
            trail_mix, spotted_dick,
            person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes,
            minibags_of_chips, kale_smoothie, joy_joy_mit_iodine, hugs_actual_physical_hugs, 
            heath_bar, healthy_fruit, glow_sticks, generic_brand_acetaminophen, 
            dental_paraphenalia, creepy_religious_comics_chick_tracts, chardonnay,
            cash_or_other_forms_of_legal_tender, broken_glow_stick, boxo_raisins, 
            bonkers_the_board_game, timestamp)) %>% 
  add_column(year = "2016", .before = 1) %>% 
  mutate(id_number = max(candy_2015_renamed$id_number) + row_number(), .before = 2)
col_removed_candy_2016 
```

2016 RENAME

```{r}
# Again checking on Mary Janes column differences 
# col_removed_candy_2016 %>% 
#   select(anonymous_brown_globs_that_come_in_black_and_orange_wrappers, mary_janes)

candy_2016_renamed <- col_removed_candy_2016 %>% 
  rename(trick_or_treating = 
           are_you_going_actually_going_trick_or_treating_yourself, 
         gender = your_gender, 
         age = how_old_are_you, 
         country = which_country_do_you_live_in, 
         state_or_prov = which_state_province_county_do_you_live_in, 
         anonymous_black_and_orange_wrapper = 
           anonymous_brown_globs_that_come_in_black_and_orange_wrappers, 
         bonkers = bonkers_the_candy, 
         restaurant_candy = 
           candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants, 
         chick_o_stick = chick_o_sticks_we_don_t_know_what_that_is, 
         gummy_bears = gummy_bears_straight_up, 
         hersheys_milk_chocolate = hershey_s_milk_chocolate, 
         licorice_black = licorice_yes_black, 
         peanut_m_ms = peanut_m_m_s, 
         party_bag_m_ms = third_party_m_ms, 
         reeses_peanut_butter_cups = reese_s_peanut_butter_cups, 
         sourpatch_kids = sourpatch_kids_i_e_abominations_of_nature, 
         sweetarts = sweet_tarts, 
         sweetums = sweetums_a_friend_to_diabetes, 
         circus_peanuts = those_odd_marshmallow_circus_peanut_things, 
         toblerone = tolberone_something_or_other)

candy_2016_renamed

```


2017 CLEAN AND ADD
Note - should have removed the "q6_" before this, but have done so in the next 
chunk 

```{r}
col_removed_candy_2017 <- janitor_candy_2017 %>%  
  select(-c(102, 104, 105, 107, 108, 110:120), 
         -c(q6_spotted_dick, 
            q6_sandwich_sized_bags_filled_with_boo_berry_crunch,
            q6_real_housewives_of_orange_county_season_9_blue_ray, 
            q6_minibags_of_chips, 
            q6_abstained_from_m_ming, 
            q6_kale_smoothie, q6_joy_joy_mit_iodine, 
            q6_hugs_actual_physical_hugs, 
            q6_heath_bar, 
            q6_healthy_fruit, 
            q6_glow_sticks, 
            q6_generic_brand_acetaminophen, 
            q6_dental_paraphenalia, 
            q6_creepy_religious_comics_chick_tracts, 
            q6_chardonnay, 
            q6_cash_or_other_forms_of_legal_tender, 
            q6_broken_glow_stick, 
            q6_boxo_raisins, 
            q6_bonkers_the_board_game, 
            internal_id)) %>% 
  add_column(year = "2017", .before = 1) %>% 
  mutate(id_number = max(candy_2016_renamed$id_number) + row_number(), .before = 2)

col_removed_candy_2017
```

2017 RENAME - get rid of "q1/2/3/4/5/6" at the start of col names
and rename to match 2015 and 16

```{r}

candy_2017_q_removed <- col_removed_candy_2017 %>% 
  rename_all(~ sub("^[q0-9]{2}_", "", 
                   make.names(names(col_removed_candy_2017))))


candy_2017_renamed <- candy_2017_q_removed %>% 
  rename(trick_or_treating = going_out, 
         state_or_prov = state_province_county_etc, 
         x100_grand_bar = `100_grand_bar`, 
         mary_janes = 
           anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes, 
         bonkers = bonkers_the_candy, 
         restaurant_candy = 
           candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants, 
         chick_o_stick = chick_o_sticks_we_don_t_know_what_that_is, 
         gummy_bears = gummy_bears_straight_up, 
         hersheys_milk_chocolate = hershey_s_milk_chocolate, 
         licorice_black = licorice_yes_black, 
         peanut_m_ms = peanut_m_m_s, 
         green_m_ms = green_party_m_ms, 
         lone_m_ms = independent_m_ms, 
         reeses_peanut_butter_cups = reese_s_peanut_butter_cups, 
         sourpatch_kids = sourpatch_kids_i_e_abominations_of_nature, 
         sweetarts = sweet_tarts, 
         sweetums = sweetums_a_friend_to_diabetes, 
         circus_peanuts = those_odd_marshmallow_circus_peanut_things, 
         toblerone = tolberone_something_or_other)
  
candy_2017_renamed
```

```{r}
view(candy_2015_renamed)
view(candy_2016_renamed)
view(candy_2017_renamed)
```



Getting an idea of people's responses: 
```{r}

distinct(candy_2015_renamed, age) 
  #(chr) 146 responses, some silly and some strange
distinct(candy_2015_renamed, trick_or_treating) 
  #(chr) yes or no (NAs in 2017)
distinct(candy_2015_renamed, starburst)


distinct(candy_2016_renamed, age) 
  # (chr) 98 incl silly/strange
distinct(candy_2016_renamed, trick_or_treating) 
  #(chr) Yes No (NAs in 2017)
distinct(candy_2016_renamed, gender) 
  # Male, Female, Other, I'd rather not say, NA
distinct(candy_2016_renamed, country) 
  #93 some silly, some e.g. USA, US, us, u.s.a. etc 
distinct(candy_2016_renamed, starburst)


distinct(candy_2017_renamed, age) 
  #(chr) 107 incl silly/strange
distinct(candy_2017_renamed, trick_or_treating)
  #(chr) Yes No and NA
distinct(candy_2017_renamed, gender) 
  #Male, Female, Other, I'd rather not say, NA
distinct(candy_2017_renamed, country) 
  #118 some silly, some e.g. USA, US, us, u.s.a. etc 
distinct(candy_2017_renamed, starburst)
```


Joining all three years by binding rows so as to keep everything

```{r}
bound_candy <- bind_rows(candy_2015_renamed, 
                         candy_2016_renamed, 
                         candy_2017_renamed)


bound_candy <- bound_candy %>% 
  relocate(country, .before = 5) %>% 
  relocate(state_or_prov, .before = 6) %>% 
  relocate(gender, .before = 7)


# view(bound_candy)
```





**QUESTION 1** 
What is the total number of candy ratings given across the three years. 
(Number of candy ratings, not the number of raters. Don’t count missing values)

ANSWER - 590,010 ratings
```{r}

total_ratings <- bound_candy %>% 
  select(-c(year, id_number, age, 
            trick_or_treating, country, 
            state_or_prov, gender)) %>%
  pivot_longer(butterfinger:take_5, 
               names_to = "candy", 
               values_to = "rating") %>% 
  filter(!is.na(rating)) %>% 
  count(n())
total_ratings

```




**AGE CLEANING**
Age is a character column with 274 values. They are numbers, NAs, and strange 
and silly values 
- changed age column to a numeric but it output strange figures, so I specified 
  integer and strange answers became NAs. 
- Seems unlikely that anyone > 100 years old is answering so I removed them

```{r}
# bound_candy %>% 
#  distinct(age)

bound_age_to_numeric <- bound_candy %>% 
  mutate(age = as.integer(age)) %>% 
  arrange(age)

# bound_age_to_numeric %>% 
# distinct(age)

bound_age_cleaning <- bound_age_to_numeric %>%
  mutate(age = if_else(age > 99, NA_integer_, age))
bound_age_cleaning
# view(bound_age_cleaning)
```

**QUESTION TWO**
What was the average age of people who are going out trick or treating?

I know the responses for trick or treating are "Yes" "No " and "NA"

ANSWER - The (rounded) average age of those going trick or treating is 35
years old (34.94897 is the unrounded value)

```{r}


bound_age_cleaning %>% 
  select(age, trick_or_treating) %>% 
  group_by(trick_or_treating) %>% 
  summarise(average_age = round(mean(age, na.rm = TRUE)))

# to just get the answer for Yes on it's own: 
bound_age_cleaning %>% 
  select(age, trick_or_treating) %>% 
  group_by(trick_or_treating) %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(average_age = round(mean(age, na.rm = TRUE)))
    
```

**QUESTION THREE**
What was the average age of people who are not going trick or treating?
ANSWER - The (rounded) average age of those not going trick or treating is 39
years old (39.10454 is the unrounded value)

```{r}
bound_age_cleaning %>% 
  select(age, trick_or_treating) %>% 
  group_by(trick_or_treating) %>% 
  summarise(average_age = round(mean(age, na.rm = TRUE)))

# to just get the answer for No on it's own: 
bound_age_cleaning %>% 
  select(age, trick_or_treating) %>% 
  group_by(trick_or_treating) %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(average_age = round(mean(age, na.rm = TRUE)))

```

**QUESTION FOUR** 
**For each of joy, despair and meh, which candy bar revived the most of these** 
**ratings?** 

As in question 1, a 2 column tibble with candy and rating was created and a count
of each distinct answer was made - firstly counting each for each response and 
each candy, then filtering to find the maximum count for each rating. 

ANSWER -  
Despair: gum that comes with baseball cards returned the most despair responses
          with 7,341
Joy:Full sized candy bars made the most people joyful with 7,589 responses.
    This seems very generic so I ran it again to remove the full sized candy
    bars and the top Joy response was: 7369 responses for reeses peanut butter
    cups
Meh: 1,570 "Meh" responses were given for lollipops


```{r}

rated_candy <- bound_candy %>% 
  select(-c(year, id_number, age, 
            trick_or_treating, country, 
            state_or_prov, gender, any_full_sized_candy_bar)) %>%
  pivot_longer(butterfinger:take_5, 
               names_to = "candy", 
               values_to = "rating")
rated_candy

rated_candy %>% 
  group_by(rating, candy) %>% 
  summarise(number_of_ratings = n()) %>% 
  filter(number_of_ratings == max(number_of_ratings))

# run again removing generic "full sized candy bars"

rated_candy <- bound_candy %>% 
  select(-c(year, id_number, age, 
            trick_or_treating, country, 
            state_or_prov, gender, any_full_sized_candy_bar)) %>%
  pivot_longer(butterfinger:take_5, 
               names_to = "candy", 
               values_to = "rating")
rated_candy

rated_candy %>% 
  group_by(rating, candy) %>% 
  summarise(number_of_ratings = n()) %>% 
  filter(number_of_ratings == max(number_of_ratings))

```


**COUNTRY CLEANING** 

Firstly getting an idea of NAs and distinct country values:

- Note that 2015 (5630 rows of 9349) has no country data... (all NA)
- 5715 rows in bound_candy has NA 

- There are 169 distinct country names including NAs, miss-spellings, and silly/ 
 unknown answers 

```{r}

bound_candy %>% 
  filter(is.na(country))

bound_candy %>% 
  distinct(country) 
```


Using stringr and regex to reduce "country" values
- For columns that were numbers or clearly fake (silly) answers, I checked 
  the state or province column to see if there was a match to a country
- I checked my work one row at a time to try and ensure I did not change 
  anything that was not meant to be changed. 
- I am sure there are faster/ better ways of doing this, but I was practicing
  different options
 

```{r}
library(stringr)

```

country = str_replace_all(country, pattern = "[ ]*[uU]+[ .!]*[sS]+[ .!]*[aA]*[ .!]*", "States")
tried this to change what was now States USA USA but it changed that to StatesStatesStates and also Australia to AStatestralia... 
```{r}
bound_country_clean <- bound_candy %>% 
  mutate(country = str_replace_all(country, pattern = "[0-9][0-9][.][0-9]", "States"),
         country = str_replace_all(country, pattern = "[3|4|5][0-9]", "States"),
         country = str_replace_all(country, pattern = "^[ ]*[uU]+[ .!]*[sS]+[ .!]*[aA]*[ .!]*", "States"),
         country = str_replace_all(country, pattern = "^[uU][nited]+\\s[sS][tT]*[aA|eE|sS][tT][eE][sS|aA]", "States"),
         country = str_replace_all(country, pattern = "^[uU][nN][iI][tT][eE][dD|sS]+\\s[sS][tT][aA][tT][eE|sS][sS]*", "States"),
         # the above doesn't work for all... i presume there are spaces somewhere
         country = str_replace_all(country, pattern = "^[uU][nN][iI][tT][sS]+\\s[sS][tT][aA][tT][eE][sS]*", "States"),
        country = str_replace_all(country, pattern = "[sS][tT][aA|eE][tT]*[eE][sS]*", "States"), 
        
         country = str_replace_all(country, pattern = "[uU]+[sS]+[aA]+", "States"),
        country = str_replace_all(country, pattern = "^[mM][uU|eE][rR]+[iI][cC|kK][aA]", "States"),
        country = str_replace_all(country, pattern = "^[aA][mM][eE][rR][iI][cC][aA]", "States"),
        country = str_replace_all(country, pattern = "^\\'[mM][uU|eE][rR][iI][cC][aA]", "States"),
        country = str_replace_all(country, pattern = "[sS][tT][aA][tT][eE][sS][!]", "States"),
        country = str_replace_all(country, pattern = "[a-zA-Z]+ [-]+ [uU][sS][aA]", "States"), 
        # the above only changed "the best one - usa" to "the best States" 
        country = str_replace_all(country, pattern = "[of] [aA][merica]", ""),
        # changes States of America to States oerica and i also have united states oerica
        country = str_replace_all(country, pattern = "cascadia", "States"),
        country = str_replace_all(country, pattern = "Narnia", "States"),
        country = str_replace_all(country, pattern = "Sub-Canadian North America... 'Merica", "States"),
        country = str_replace_all(country, pattern = "Trumpistan", "States"),
        country = str_replace_all(country, pattern = "The republic of Cascadia", "States"),
        country = str_replace_all(country, pattern = "unhinged ", ""),
        country = str_replace_all(country, pattern = "North Carolina", "States"),
        country = str_replace_all(country, pattern = "Pittsburgh", "States"),
        country = str_replace_all(country, pattern = "New York", "States"),
        country = str_replace_all(country, pattern = "Ahem....Amerca", "States"),
        country = str_replace_all(country, pattern = "UD", "States"),  
        country = str_replace_all(country, pattern = "New Jersey", "States"),
        country = str_replace_all(country, pattern = "Alaska", "States"),
        country = str_replace_all(country, pattern = "N. America", "States"),
        country = str_replace_all(country, pattern = "I don't know anymore", "States"),
        country = str_replace_all(country, pattern = "!.*", ""),
        country = str_replace_all(country, pattern = "States[sS|dD|aA]", "States"),
        country = str_replace_all(country, pattern = "States oerica", "States"),
        country = str_replace_all(country, pattern = "Statesof A", "States"),
        country = str_replace_all(country, pattern = "United States", "States"),
        country = str_replace_all(country, pattern = "The United States", "States"),
        country = str_replace_all(country, pattern = "The States", "States"),
        country = str_replace_all(country, pattern = "Statestates States", "States"),
        country = str_replace_all(country, pattern = "Statestates States States", "States"),
        country = str_replace_all(country, pattern = "Statestates", "States"),
        country = str_replace_all(country, pattern = "States States", "States"),
        country = str_replace_all(country, pattern = "^[uU]+[.]*[kK]+[.]*", "United Kingdom"),
        country = str_replace_all(country, pattern = "^[uU][nited]+\\s[kK][iI][nN][dD][oO][mM]", "United Kingdom"),
        country = str_replace_all(country, pattern = "^[uU][nN][iI][tT][eE][dD]+\\s[kK][iI][nN][gG][dD][oO][mM]", "United Kingdom"),
        country = str_replace_all(country, pattern = "^[eE][nN][gG|dD][lL][aA][nN][dD]", "United Kingdom"),
        country = str_replace_all(country, pattern = "Scotland", "United Kingdom"),
        country = str_replace_all(country, pattern = "^[cC][a-zA-Z]{5}", "Canada"),
        country = str_replace_all(country, pattern = "^[cC]+[aA]+[nN]+[aA]+[dD]+[aA]+[aA|iI|rR]*[aA|nN]*[iI]*[aA]*", "Canada"),
        country = str_replace_all(country, pattern = "Canae", "Canada"),
        country = str_replace_all(country, pattern = "^Canada`", "Canada"),
        country = str_replace_all(country, pattern = "soviet canuckistan`", "Canada")
      
        )

bound_country_clean %>% 
  distinct(country)
view(bound_country_clean)
```            

Still and issue:
# States? Hard to tell anymore..
# one "United States" not changing (probably spaces issue??)
# States (I think but it's an election year so who can really tell)
# I pretend to be from Canada, but I am really from the United States.
# there isn't one for old men
# one of ones		(previously "one of the good ones"?)
# The Yoo Essaayyyyyy
# this one
# neverland 
# somewhere
# god's country 
# EUA
# See above
# Not the or Canada (previously not the US or canada)
# Denial 
# Earth 
# insanity lately 
# A 
# Can 
# Atlantis
# Narnia
# 1 
# subscribe to dm4uz3 on youtube		
# Fear and Loathing


