---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
```

```{r}
candy_2015 <- read_excel("../raw_data/boing-boing-candy-2015.xlsx")
candy_2016 <- read_excel("../raw_data/boing-boing-candy-2016.xlsx")
candy_2017 <- read_excel("../raw_data/boing-boing-candy-2017.xlsx")

here::here()
```

From glimpse 
```{r}
head(candy_2015)
glimpse(candy_2016)
glimpse(candy_2017)
```



```{r}
library(janitor)

```

```{r}

janitor_candy_2015 <- janitor::clean_names(candy_2015)
janitor_candy_2015

janitor_candy_2016 <- janitor::clean_names(candy_2016)
janitor_candy_2016

janitor_candy_2017 <- janitor::clean_names(candy_2017)
janitor_candy_2017
```


just looking at who is reporting back about these ones... 
```{r}

janitor_candy_2017 %>% 
  select(q3_age, q2_gender, q6_independent_m_ms, q6_green_party_m_ms)

janitor_candy_2016 %>% 
  select(york_peppermint_patties_ignore)
#nobody 
```


2015 clean - remove columns that aren't candy (see readme) I went from bottom to
top to check index as i went
Also adding a "year" column and removing "timestamp" for the join (see if can work 
out extracting year from timestamp and moving over later). Doing this after so that 
column index isn't ruined 
```{r}
names(janitor_candy_2015)

col_removed_candy_2015 <- janitor_candy_2015 %>% 
  select(-c(116:124), -c(97:113), -c(93:95), -c(90, 91), -c(peterson_brand_sidewalk_chalk, spotted_dick, mint_leaves, joy_joy_mit_iodine, minibags_of_chips, lapel_pins, kale_smoothie, hugs_actual_physical_hugs, heath_bar, healthy_fruit, creepy_religious_comics_chick_tracts, broken_glow_stick, glow_sticks, generic_brand_acetaminophen, dental_paraphenalia, cash_or_other_forms_of_legal_tender, vials_of_pure_high_fructose_corn_syrup_for_main_lining_into_your_vein, box_o_raisins, timestamp)) %>% 
  add_column(year = "2015", .before = 1)
  

col_removed_candy_2015
#view(col_removed_candy_2015)


```

2015 clean - renaming columns so they match other years
considered merging anonymous brown globs with mary janes (in 2017 it is "anon 
brown... aka Mary Janes") but looking at responses, they don't match up so I'll 
leave as anonymous_black_and_orange_wrappers...
```{r}
col_removed_candy_2015 %>% 
  select(anonymous_brown_globs_that_come_in_black_and_orange_wrappers, mary_janes)

candy_2015_renamed <- col_removed_candy_2015 %>% 
  rename(age = how_old_are_you, trick_or_treating = are_you_going_actually_going_trick_or_treating_yourself, anonymous_black_and_orange_wrapper = anonymous_brown_globs_that_come_in_black_and_orange_wrappers, brach_not_including_candy_corn = brach_products_not_including_candy_corn, restaurant_candy = candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants, hersheys_dark_chocolate = dark_chocolate_hershey, gummy_bears = gummy_bears_straight_up, hersheys_kissables = hershey_s_kissables, hersheys_milk_chocolate = hershey_s_milk_chocolate, licorice_black = licorice, reeses_peanut_butter_cups = reese_s_peanut_butter_cups, toblerone = tolberone_something_or_other, peanut_m_ms = peanut_m_m_s, chick_o_stick = chick_o_sticks_we_don_t_know_what_that_is, circus_peanuts = those_odd_marshmallow_circus_peanut_things, sea_salt_chocolate = sea_salt_flavored_stuff_probably_chocolate_since_this_is_the_it_flavor_of_the_year)

```


2016 clean - remove unnecessary columns - bottom to top
Also adding a year column for the binding of rows
```{r}
col_removed_candy_2016 <- janitor_candy_2016 %>% 

  select(-c(104, 105, 107:123), -c(vicodin, vials_of_pure_high_fructose_corn_syrup_for_main_lining_into_your_vein, trail_mix, spotted_dick, person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes, minibags_of_chips, kale_smoothie, joy_joy_mit_iodine, hugs_actual_physical_hugs, heath_bar, healthy_fruit, glow_sticks, generic_brand_acetaminophen, dental_paraphenalia, creepy_religious_comics_chick_tracts, chardonnay, cash_or_other_forms_of_legal_tender, broken_glow_stick, boxo_raisins, bonkers_the_board_game, timestamp)) %>% 
  add_column(year = "2016", .before = 1)


```

2016 - rename so they match other sheets
considered merging anonymous brown globs with mary janes (in 2017 it is "anon 
brown... aka Mary Janes") but looking at responses, they don't match up so I'll 
leave as anonymous_black_and_orange_wrappers...
```{r}

col_removed_candy_2016 %>% 
  select(anonymous_brown_globs_that_come_in_black_and_orange_wrappers, mary_janes)

candy_2016_renamed <- col_removed_candy_2016 %>% 
  rename(trick_or_treating = are_you_going_actually_going_trick_or_treating_yourself, gender = your_gender, age = how_old_are_you, country = which_country_do_you_live_in, state_or_prov = which_state_province_county_do_you_live_in, anonymous_black_and_orange_wrapper = anonymous_brown_globs_that_come_in_black_and_orange_wrappers, bonkers = bonkers_the_candy, restaurant_candy = candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants, chick_o_stick = chick_o_sticks_we_don_t_know_what_that_is, gummy_bears = gummy_bears_straight_up, hersheys_milk_chocolate = hershey_s_milk_chocolate, licorice_black = licorice_yes_black, peanut_m_ms = peanut_m_m_s, party_bag_m_ms = third_party_m_ms, reeses_peanut_butter_cups = reese_s_peanut_butter_cups, sourpatch_kids = sourpatch_kids_i_e_abominations_of_nature, sweetarts = sweet_tarts, sweetums = sweetums_a_friend_to_diabetes, circus_peanuts = those_odd_marshmallow_circus_peanut_things, toblerone = tolberone_something_or_other)

candy_2016_renamed

```


2017 clean - remove those not candy and internal id as others don't have
Also adding a year column for the join
```{r}
col_removed_candy_2017 <- janitor_candy_2017 %>%  
  select(-c(102, 104, 105, 107, 108, 110:120), -c(q6_spotted_dick, q6_sandwich_sized_bags_filled_with_boo_berry_crunch, q6_real_housewives_of_orange_county_season_9_blue_ray, q6_minibags_of_chips, q6_abstained_from_m_ming, q6_kale_smoothie, q6_joy_joy_mit_iodine, q6_hugs_actual_physical_hugs, q6_heath_bar, q6_healthy_fruit, q6_glow_sticks, q6_generic_brand_acetaminophen, q6_dental_paraphenalia, q6_creepy_religious_comics_chick_tracts, q6_chardonnay, q6_cash_or_other_forms_of_legal_tender, q6_broken_glow_stick, q6_boxo_raisins, q6_bonkers_the_board_game, internal_id)) %>% 
  add_column(year = "2017", .before = 1)

col_removed_candy_2017
```

2017 - rename - get rid of q1/2/3/4/5/6 at the start of col names
and rename to match 2015 and 16

```{r}

candy_2017_q_removed <- col_removed_candy_2017 %>% 
  rename_all(~ sub("^[a-z0-9]{2}_", "", 
                   make.names(names(col_removed_candy_2017))))

candy_2017_renamed <- candy_2017_q_removed %>% 
  rename(trick_or_treating = going_out, state_or_prov = state_province_county_etc, x100_grand_bar = `100_grand_bar`, mary_janes = anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes, bonkers = bonkers_the_candy, restaurant_candy = candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants, chick_o_stick = chick_o_sticks_we_don_t_know_what_that_is, gummy_bears = gummy_bears_straight_up, hersheys_milk_chocolate = hershey_s_milk_chocolate, licorice_black = licorice_yes_black, peanut_m_ms = peanut_m_m_s, green_m_ms = green_party_m_ms, lone_m_ms = independent_m_ms, reeses_peanut_butter_cups = reese_s_peanut_butter_cups, sourpatch_kids = sourpatch_kids_i_e_abominations_of_nature, sweetarts = sweet_tarts, sweetums = sweetums_a_friend_to_diabetes, circus_peanuts = those_odd_marshmallow_circus_peanut_things, toblerone = tolberone_something_or_other)
  
candy_2017_renamed
```

```{r}
view(candy_2015_renamed)
view(candy_2016_renamed)
view(candy_2017_renamed)
```



Getting an idea of people's responses: 
```{r}

distinct(candy_2015_renamed, age) #(chr...) 146 responses, some silly and some strange
distinct(candy_2015_renamed, trick_or_treating) # yes or no (chr)
distinct(candy_2015_renamed, starburst)


distinct(candy_2016_renamed, age) #chr 98 incl silly/strange
distinct(candy_2016_renamed, trick_or_treating) #Yes No (chr)
distinct(candy_2016_renamed, gender) #Male, Female, Other, I'd rather not say, NA
distinct(candy_2016_renamed, country) #93 some silly, some e.g. USA, US, us, u.s.a. etc 
distinct(candy_2016_renamed, starburst)


distinct(candy_2017_renamed, age) #chr 107 incl silly/strange
distinct(candy_2017_renamed, trick_or_treating) #Yes No and NA(chr)
distinct(candy_2017_renamed, gender) #Male, Female, Other, I'd rather not say, NA
distinct(candy_2017_renamed, country) #118 some silly, some e.g. USA, US, us, u.s.a. etc 
distinct(candy_2017_renamed, starburst)
```

Join all three 
```{r}
bound_candy <- bind_rows(candy_2015_renamed, candy_2016_renamed, candy_2017_renamed)

bound_candy <- bound_candy %>% 
  relocate(country, .before = 4) %>% 
  relocate(state_or_prov, .before = 5) %>% 
  relocate(gender, .before = 6)


view(bound_candy)
```



---------------- COUNTRY CLEANING -------------------------- 
Note that 2015 (5630 rows of 9349) has no country data... this seems to be 
American so maybe change NAs to USA at later date(??)

```{r}
bound_candy %>% 
  distinct(country)
# 169 country variations 

bound_candy %>% 
  filter(is.na(country))
# 5715 rows in country are NA
```

```{r}
#just to see that there are the right amount of NAs 
#country_cleaning <- bound_candy %>% 
 # select(country, state_or_prov) %>% 
  #filter(!is.na(country))
# 3634 rows remain with NAs gone... 
#view(country_cleaning)

# first... remove punctuation
library(stringr)

#below is deff wrong 
#country_clean <- bound_candy %>% 
 # str_replace_all("us" | "US" | "united states" | "usa" | "uSA" | "United States of America" | "United States" | "USSA" | "U.S.A." | "USA!", "USA")
#country_clean

#candy_2016_renamed %>% 
#  mutate(country = str_replace_all(country, pattern = "\\.\\!", "")
#view(candy_2016_renamed)
  

country_2016 <- candy_2016_renamed %>% 
  mutate(country = str_replace_all(country, pattern = "^[uU]+[sS]+[aA]+", "America"),
         country = str_replace_all(country, pattern = "^[uU]+\\.+[sS]+\\.+[aA]+\\.", "America"),
         country = str_replace_all(country, pattern = "^[Uu]+[Ss]+[Aa]+[!{1, }]", "America")
         #the ! arguement isn't working!
        # country = str_replace_all(country, pattern = "^[uU]+[sS]+[aA]+[:punct:]+", "America")
         )
view(country_2016)


str_replace_all(my_string, "[^[:alnum:]]", "") 

```

mutate(col name, 
      .fns = ~str_remove_all(., 
      "JUV|AD|IMM|SUB|PL*|[0-9]|", 
      .names = 'fixed_{.col}')
      )
      
      

AGE QUESTION 
```{r}
#bound_candy %>% 
 # distinct(age)

#274 different ages (chr)

#bound_candy <- bound_candy %>% 
 # mutate(age = as.double(age))
#need to remove all non-numbers and NAs first 

#age_focus <- bound_candy %>% 
 # select(age) %>% 
  #filter(age < "90") %>% 
  #arrange(desc(age))#
#view(age_focus) # still have 0.62, 142, 1e+18 etc in there...

```

library(stringr)

year_pattern <- "^[0-9]{4}"

col_removed_candy_2015 %>% 
  select(timestamp) %>% 
  str_extract(year_pattern)